---
title: "Plate Correction"
author: "Guillaume Blanchet"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### R package

```{r, message=FALSE}
# For spatial analyses
library(plateCorrect)

# plot figure
library(fields)
# Nice colour palettes
library(dichromat)
```

### Read test data
```{r}
data(plate)
```

#### Organise data in a matrix format

Note that to organise the data using the code  below, it is essential to have a references columns defining how the data in the plate should be ordered.

```{r}
# Extract plate row and column names
rowName <- substr(plate[,1],1,2)
colName <- substr(plate[,1],3,4)

# Convert these names into numbers
rowFac <- as.factor(rowName)
colFac <- as.factor(colName)

rowRefNum <- as.numeric(rowFac)
colRefNum <- as.numeric(colFac)

# Extract number of rows and columns in the plate
nRows <- nlevels(rowFac)
nCols <- nlevels(colFac)

# Build the matrix
mat1536 <- matrix(NA, nrow = nRows, ncol = nCols)

for(i in 1: length(rowRefNum)){
  mat1536[rowRefNum[i],colRefNum[i]] <- plate[i,2]
}
```

### Standardize data

```{r}
matStand <- standardize(mat1536, FUN = scale)
```

### Correction
#### Row-Column

# Write the math of the methods... to be done later

```{r}
matRC.25_.75 <- correctRC(matStand, lower = 0.25, upper = 0.75)
matRC.4_.6 <- correctRC(matStand, lower = 0.4, upper = 0.6)
```

#### Autocorrelation-based correction

The model used to perform this analysis is 

$$y = \beta_0 + u$$
where

$$u = \lambda \mathbf{W} u + \varepsilon.$$

In the previous equation, $\beta_0$ is an intercept and $\lambda$ is a parameter that weight the importance of the spatial structure considered, defined by $\mathbf{W}u$, and $\varepsilon$ is an error term. Note that $\mathbf{W}u$ is a way to define how the spatial autocorrelation $u$ is structured. 

```{r}
matAuto <- correctAuto(matStand)
matAuto_Queen <- correctAuto(matStand, diagWeight = 1)
matAuto_Rook <- correctAuto(matStand, diagWeight = 0)
```

###### Plot the result for comparison
```{r}
# Color for figures
couleur <- colorRampPalette(colorschemes$Categorical.12)(100)

# Plot Raw
image.plot(1:48,1:32,
      t(matStand)[,nrow(matStand):1],
      col = couleur,
      zlim = range(matStand,
                   matAuto,
                   matAuto_Queen,
                   matAuto_Rook),
      main = "Raw data")

# Plot prediction (function default)
image.plot(1:48,1:32,
      t(matAuto)[,nrow(matAuto):1],
      col = couleur,
      zlim = range(matStand,
                   matAuto,
                   matAuto_Queen,
                   matAuto_Rook),
      main = "Function default")

# Plot prediction (Queen)
image.plot(1:48,1:32,
      t(matAuto_Queen)[,nrow(matAuto_Queen):1],
      col = couleur,
      zlim = range(matStand,
                   matAuto,
                   matAuto_Queen,
                   matAuto_Rook),
      main = "Queen")

# Plot prediction (Rook)
image.plot(1:48,1:32,
      t(matAuto_Rook)[,nrow(matAuto_Rook):1],
      col = couleur,
      zlim = range(matStand,
                   matAuto,
                   matAuto_Queen,
                   matAuto_Rook),
      main = "Rook")

# Plot prediction (Row-columns standardization)
image.plot(1:48,1:32,
      t(matRC.25_.75)[,nrow(matRC.25_.75):1],
      col = couleur,
      zlim = range(matRC.25_.75),
      main = "RC")

# Plot prediction (Row-columns standardization)
image.plot(1:48,1:32,
      t(matRC.4_.6)[,nrow(matRC.4_.6):1],
      col = couleur,
      zlim = range(matRC.4_.6),
      main = "RC")
```
